\documentclass{sciposter}
\usepackage[dvipsnames,usenames,svgnames,table,x11names, rgb, html]{xcolor} 
\usepackage{lipsum}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[german]{babel}
\usepackage{geometry}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{wrapfig}
\usepackage{gensymb}
\usepackage[utf8]{inputenc}
\usepackage{empheq}
\usepackage{mathpazo}
\usepackage[T1]{fontenc}
\renewcommand{\familydefault}{\rmdefault}
\usepackage{eulervm}

% for nice tableas
\usepackage{booktabs}

% so we have more space between matrix entries
\setlength{\arraycolsep}{12pt}

\geometry{
 landscape,
 a1paper,
 left=5mm,
 right=50mm,
 top=5mm,
 bottom=50mm,
 }


\usepackage{todonotes}

\newcommand{\nnz}{d\operatorname{nnz}}
\newcommand{\TODO}[1]{\todo[inline, color=red!40]{#1}}
\renewcommand{\vec}[1]{\mathbf{#1}}
%BEGIN LISTINGDEF



\usepackage{listings}
\usepackage{inconsolata}


\usepackage{mdframed}

\newenvironment{method}[1]{\begin{mdframed}[backgroundcolor=blue!10,innertopmargin=15pt, innerbottommargin=15pt, nobreak=true]
		\textbf{#1 }
	}
	{ 
	\end{mdframed}
}

\newenvironment{important}{\begin{mdframed}[backgroundcolor=red!50,innertopmargin=15pt, innerbottommargin=15pt, nobreak=true]
		\Large
	}
	{ 
	\end{mdframed}
}

\newenvironment{trick}[1]{\begin{mdframed}[backgroundcolor=PineGreen!50,innertopmargin=15pt, innerbottommargin=15pt, nobreak=true]
			\textbf{#1 }
	}
	{ 
	\end{mdframed}
}


\definecolor{background}{HTML}{FAFAFA}
\definecolor{comment}{HTML}{ABB0B6}
\definecolor{keywords}{HTML}{55B4D4}
\definecolor{basicStyle}{HTML}{6C7680}
\definecolor{variable}{HTML}{001080}
\definecolor{string}{HTML}{86B300}

\newcommand{\norm}[1]{\left\lVert#1\right\rVert}


\lstset{
	% How/what to match
	sensitive=true,
	% Border (above and below)
	frame=single,
	% Extra margin on line (align with paragraph)
	xleftmargin=\parindent,
	% Put extra space under caption
	belowcaptionskip=1\baselineskip,
	% Colors
	backgroundcolor=\color{background},
	basicstyle=\color{basicStyle}\ttfamily,
	keywordstyle=\color{keywords},
	commentstyle=\color{comment},
	stringstyle=\color{string},
	numberstyle=\color{sviolet},
	identifierstyle=\color{variable},
	% Break long lines into multiple lines?
	breaklines=true,
	% Show a character for spaces?
	showstringspaces=false,
	tabsize=2
}

%END LISTINGDEF


\newcommand*\widefbox[1]{\fbox{\hspace{2em}#1\hspace{2em}}}


\newlength\dlf  % Define a new measure, dlf
\newcommand\alignedbox[2]{
% Argument #1 = before & if there were no box (lhs)
% Argument #2 = after & if there were no box (rhs)
&  % Alignment sign of the line
{
\settowidth\dlf{$\displaystyle #1$}
    % The width of \dlf is the width of the lhs, with a displaystyle font
\addtolength\dlf{\fboxsep+\fboxrule}
    % Add to it the distance to the box, and the width of the line of the box
\hspace{-\dlf}
    % Move everything dlf units to the left, so that & #1 #2 is aligned under #1 & #2
\boxed{#1 #2}
    % Put a box around lhs and rhs
}
}
\usepackage{graphicx,url}

%BEGIN TITLE
\title{\huge{Case Studies from Practice}}

% if you want to set an author replace the next line with \author{your name here}
\author{David Bohner}



\begin{document}
\fontfamily{lmss}\selectfont
\maketitle

\begin{multicols}{3}

\section{General Pointers}
\subsection{General Form}

\begin{mdframed}
\begin{itemize}
	\item \underline{Always} compare your suggestion to the \textit{'Do Nothing'} option.
	\begin{itemize}
		\item e.g. increased maintenance costs for outdated software when considering spending to re-develop it in-house.
	\end{itemize}
	\item Present several alternatives.
	\item Explain and Defend: Is it an actual problem worth fixing, an opportunity worth pursuing?
	\begin{itemize}
		\item Opportunity, Risk, Cost?
	\end{itemize}
\end{itemize}
\end{mdframed}

\begin{important}{}
Reference the \texttt{Problem Statement Template!}
\end{important}

\subsection{Interest of Managers}

\begin{mdframed}
\begin{itemize}
	\item \textbf{(Senior) Management} focuses on:
	\begin{itemize}
		\item Higher Revenue
		\item Lower Costs
		\item Fewer Risks
	\end{itemize}
\end{itemize}
\end{mdframed}

\section{Applied Analytic Methods}
\subsection{Software Changes / Upgrades}
\subsubsection{Open-Source Software}

\begin{mdframed}
\begin{itemize}
\item Are there available open-source variants of the software you need?
	\begin{itemize}
		\item Is it viable to use?
		\begin{itemize}
			\item Available Features
			\item Stability
			\item 3rd-Party Support (unlikely)
			\item Speed of feature updates, security patches?
			\item Continued update/support (longevity)?
			\item Who funds the organization?
			\item What are the interests of the organization?
		\end{itemize}
	\end{itemize}
	
\item Should we use OSS?
\item Should we open-source our own software?
\item What is the impact if \textit{a competitor} open-source their software?
\item Should we sponsor/donate to OSS projects? What benefit do we receive from this?
\end{itemize}
\end{mdframed}


\end{multicols}
\end{document}