\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
%\usepackage{tikz}
%\usepackage{graphicx}
%\usetikzlibrary{positioning}
\usepackage{tabularx}

\usepackage{hyperref}
\usepackage{color}
\hypersetup{
	colorlinks,
	filecolor=black,
	linkcolor=black,
	urlcolor=black
}
\usepackage{listings}
%\usepackage[table]{xcolor}
%\setlength{\parindent}{0cm}

%TODO : 

\author{David Bohner}
\title{Systems Programming and Computer Architecture \\ Summary}
\date{HS 2019}

\begin{document}
\maketitle
\tableofcontents
\newpage
\section{Definitions}

\section{5 'Realities'}
\begin{itemize}
	\item \textbf{int}s and \textbf{float}s are \textbf{not} numbers.
		\begin{itemize}
			\item $x^2 \geq 0?$
			\begin{itemize}
				\item Floats?: Yes!
				\item Ints?: Overflows!
			\end{itemize}
			\item $(x+y)+z = x+(y+z)?$
			\begin{itemize}
				\item Ints?: Yes!
				\item Floats?: Not always!
			\end{itemize}
		\end{itemize}
	\item You have to know assembly
		\begin{itemize}
			\item To understand:
			\begin{itemize}
				\item \textbf{bugs}
				\item program \textbf{performance}
				\item implement \textbf{system software}
				\item create/fight \textbf{malware}
			\end{itemize}
		\end{itemize}
	\item Memory matters. (RAM is not a realistic abstraction).
		\begin{itemize}
			\item Memory:
			\begin{itemize}
				\item is \textbf{not unbounded}
				\item \textbf{performance} is \textbf{not uniform}
				\item is \textbf{typed} (= different behaviours)
			\end{itemize}
		\end{itemize}
	\item There's much more to performance than asymptotic complexity.
		\begin{itemize}
			\item Memory fetch times (access patterns) much more important nowadays
		\end{itemize}
	\item Computers don't just execute programs / calculate values.
		\begin{itemize}
			\item e.g. I/O
		\end{itemize}
\end{itemize}

\section{Crash-Course to C}
\subsection{Control Flow}
\begin{itemize}
	\item Functions
		\begin{itemize}
			\item return (..)
		\end{itemize}
	\item Loops
		\begin{itemize}
			\item for (..; ..; ..)
			\item do .. while (..)
			\item while (..) ..
		\end{itemize}
	\item Conditionals
		\begin{itemize}
			\item if (..) then .. else ..
			\item switch (..) case .. : ..; default ..
		\end{itemize}
	\item Jumps
		\begin{itemize}
			\item break, continue
			\item goto ..
		\end{itemize}
	\item I/O
		\begin{itemize}
			\item printf()
		\end{itemize}
\end{itemize}

\subsection{Operators}
\begin{itemize}
	\item Operators and \textbf{precedence}
	\item Assignment operators
	\item Post/pre inc/decrement
	\item Casting
\end{itemize}

\subsection{Arrays/Strings}
\begin{itemize}
	\item Arrays of basic types
	\item Multidimensional arrays
	\item Initializers
	\item Strings
	\begin{itemize}
		\item Arrays of ASCII characters
		\item Null-terminated (str[5] = \texttt{\{'n', 'u', 'l', 'l', '\textbackslash 0'\}})
	\end{itemize}
	\item String library
\end{itemize}
\end{document}